---
title: "Graphics with plot"
author: 'Yuqing'
date: '10/22/2018'
output: html_document
---

# Learning plot 

We are learning plot. It's going to be amazing.

``` {r steup, include=False}
knitr:: opts_chunk$set(echo=TRUE)
```

## Load Packages

```{r packages}

# install.packages('tidyverse')
## tidyverse is a 'packages of packages ' and contians many of the packages we will use in the work shop , incluseding ggplt

# load installed pakages using 'labrafy()'
library(tidyverse)

```

## Load NP data

# install and load tidyverse
install.packages(“tidyverse”) # note on Windows and linux use double quotes
library(tidyverse)

#National Parks in California
ca <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/ca.csv") 

#Acadia National Park
acadia <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/acadia.csv")

#Southeast US National Parks
se <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/se.csv")

#2016 Visitation for all Pacific West National Parks
visit_16 <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/visit_16.csv")

#All Nationally designated sites in Massachusetts
mass <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/mass.csv")



start with the Califoria ('Ca') dataset

``` {r}
head (ca)

```



