---
title: "tidydata"
output: html_document
---
```{r}

library(tidyverse)


## read in gapminder data in wide format
gap_wide <- readr::read_csv('https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder_wide.csv')


```
## tidyr:: gatere()

```{r}

gap_long <- gap_wide %>%
  gather(key=obstype_year,
         value=obs_values,
         -continent, -country)

```

## tidyr:: separate()
```{r}

gap_long <- gap_wide %>%
  gather(key=obstype_year,
         value=obs_values, 
         -continent, -country)  %>% 
  separate(col=obstype_year,
           into=c('obstype','year'),
           sep='_',
           convert = TRUE)

```

## plot long data
```{r}
canada_df <- gap_long  %>% 
  filter(country=='Canada',
         obstype=='lifeExp')

ggplot(data=canada_df,aes(x=year,y=obs_values))+
  geom_line()
  

```


# plot the life expectancy of all conutries inthe americans
```{r}
life_df <- gap_long  %>% 
  filter(obstype=='lifeExp',
         continent=='Americas')

ggplot(data = life_df,aes(x=year,y=obs_values,color=country))+
  geom_line()



```



Using gap_long, calculate and plot the the mean life expectancy for each continent over time from 1982 to 2007. Give your plot a title and assign x and y labels. Hint: use the dplyr::group_by() and dplyr::summarize() functions.
STOP: Knit the R Markdown file and sync to Github (pull, stage, commit, push)


```{r}


 life_ev <- gap_long %>% 
  group_by(continent,obstype)  %>% 
  summarize(mean_lf=mean(obs_values))

    
  cont<- gap_long %>% 
    filter(obstype=='lifeExp',
           year>1980 & year<2006) %>% 
    group_by(continent,year) %>% 
    summarize(mean_le=mean(obs_values))
  
ggplot(data=cont,aes(x=year,y=mean_le, color=continent))+
  geom_line()+
  labs(title='Mean life Expectancy',
       x='year',
       y='Age (years)')+
  theme_bw()


```

## spread
```{r}

gap_normal<- gap_long %>% 
  spread(key = obstype,
         value=obs_values)

```



## unite
```{r}

gap_uni <- gap_long %>% 
  unite(obstype_year,obstype,year) %>% 
  spread(key=obstype_year,
         value=obs_values)


gap_wide_new <- gap_long  %>% 
  unite(col=var_names,obstype,year,sep='_') %>% 
  spread(key=var_names,
         value=obs_values)


```


gap_long <- gap_wide %>%
  gather(key=obstype_year,
         value=obs_values, 
         -continent, -country)  %>% 
  separate(col=obstype_year,
           into=c('obstype','year'),
           sep='_',
           convert = TRUE)







